---
title: "Assignment 2"
output:
  html_document:
    df_print: paged
date: "2026-04-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r cars}
library(readr)
library(table1) # package for generating tables
library(tidyverse)
library(ggplot2)
library(sjPlot) #package for plotting probabilities of outcome based on LR

#load data
cohort <- read_csv("/Users/meri/Documents/Postdoc/T32/HPH203/cohort.csv")
#inspect data
head(cohort)

#specify levels and labels for factor variables
cohort <- cohort %>%
  mutate(
    female = factor(female, 
                    levels = c(0,1), 
                    labels = c("Male", "Female")),
    smoke = factor(smoke, 
                    levels = c(0,1), 
                    labels = c("No smoking", "Smoking")),
    cardiac = factor(cardiac, 
                    levels = c(0,1), 
                    labels = c("No cardiac outcome", "Cardiac outcome")),
    
  )

```
```{r}
## create table of describing data (stratified by chosen outcome variable "cardiac")
table1(~age+ factor(female)+ factor(smoke)+ cost| cardiac, data=cohort)
```




```{r}
## Logistic regression to look at odds of cardiac outcome in males vs females adjusted for age, smoking status and money spent on cigarettes as proxy of amount smoked

## Step 1: unadjusted LR for each variable
m_sex <- glm(cardiac~female, data=cohort, family="binomial")
summary(m_sex)
round(exp(cbind(OR = coef(m_sex), confint(m_sex))),2)

m_age <- glm(cardiac~age, data=cohort, family="binomial")
summary(m_age)
round(exp(cbind(OR = coef(m_age), confint(m_age))),2)

m_smoke <- glm(cardiac~smoke, data=cohort, family="binomial")
summary(m_smoke)
round(exp(cbind(OR = coef(m_smoke), confint(m_smoke))),2)

m_cost <- glm(cardiac~cost, data=cohort, family="binomial")
summary(m_cost)
round(exp(cbind(OR = coef(m_cost), confint(m_cost))),2)
```


```{r}
## Step 2: adjusted LR with all variables in the model
m2 <- glm(cardiac~female+ smoke +age +cost, data=cohort, family="binomial")

summary(m2)
round(exp(cbind(OR = coef(m2), confint(m2))),2)
# The coefficients and odds ratios are unexpected considering the positive association seen in unadjusted models
# Cost could be a consequence of cardiac outcome rather than a predictor, which would distort the analysis
```


```{r}
## Stpe 3: final full LR model adjusted for sex, smoking and age, but leaving out cost
m3 <- glm(cardiac~female+ smoke +age, data=cohort, family="binomial")

summary(m3)
round(exp(cbind(OR = coef(m3), confint(m3))),2)
```

```{r}
## Plot the predicted probabilities of a cardiac outcome based on age, sex, and smoking status

plot_model(m3,
  type = "pred",
  terms = c("age", "smoke", "female")
) +
  labs(y = "Prob(cardiac)") 
```




## Summary of findings 
This analysis explored demographic and behavioral risk factors for experiencing a cardiac outcome in a cohort of 5000 individuals. Individuals who experienced cardiac outcomes (N=275) were more likely to be male (88%) than female (12%) and have a history of smoking than individuals without cardiac outcomes (48.4% vs.10.9% smokers, respectively). Individuals with cardiac outcomes were also older than individuals who did not experience cardiac outcomes (mean (SD) 44.9 (15.7) vs. 46.4 (16.2)). In logistic regression analyses, older age, male sex and history of smoking emerged as a significant predictors of cardiac outcomes in both unadjusted and adjusted analyses in indicating independent contributions to cardiac risk.
```{r}

```



